---
title: "Healthcare Capacity in New York State"
author: Eileen Ternize
format:
  html:
    theme: yeti
execute:
  echo: false
---

## Data

The dataset combines data from the Centers for Disease Control and Prevention’s [Behavioral Risk Factor Surveillance System (BRFSS)](https://www.cdc.gov/brfss/smart/smart_2017.html), a collection of health-related surveys conducted each year of more than 400,000 Americans, and the [Kaiser Family Foundation](https://khn.org/news/as-coronavirus-spreads-widely-millions-of-older-americans-live-in-counties-with-no-icu-beds/#lookup) to show the number of people who are at high risk of becoming seriously ill from COVID-19 per ICU bed in each metropolitan area, micropolitan area or metropolitan division for which we have data.

*\*Data was found on github by jayb, covid-geography*

```{r}
dat <- read_csv(here::here("data/raw/data.csv"))
```

```{r}
#| warning: false
#| message: false
library(here)
library(tidyverse)
library(gtsummary)
library(dplyr)
library(here)

mmsa_icu_beds <- read_csv(here::here( "mmsa-icu-beds.csv"))
ny_mmsas <- mmsa_icu_beds %>%
  filter(str_detect(MMSA, "NY"))
```

The cleaned dataset has `r nrow(ny_mmsas)` MMSAs across New York state with complete data on total high-risk individuals and hospital counts.

## **Summary**

```{r}
#| echo: false
#| label: tbl-one
#| tbl-cap: "Healthcare Capacity across metropolitain cities in New York"
tbl_summary(ny_mmsas,
  by = MMSA,
  include = c(hospitals,high_risk_per_hospital, total_at_risk),
  missing = "no",
  type = list(hospitals ~ "continuous",
              high_risk_per_hospital ~ "continuous",
              total_at_risk ~ "continuous"),
statistic = all_continuous() ~ "{mean}",
  label = list(high_risk_per_hospital ~ "High risk individuals per hospital",
               hospitals ~ "Hospitals in area",
               total_at_risk ~ "Total number of high risk individuals")) |>
add_overall(col_label = "**Mean**") |>
bold_labels() |>
remove_footnote_header() |>
modify_header(label = "**Variable**") |>
modify_caption("Healthcare Capacity across metropolitain cities in New York") 

table1 <- tbl_summary(ny_mmsas,
  by = MMSA,
  include = c(hospitals,high_risk_per_hospital, total_at_risk),
  missing = "no",
  type = list(hospitals ~ "continuous",
              high_risk_per_hospital ~ "continuous",
              total_at_risk ~ "continuous"),
statistic = all_continuous() ~ "{mean}",
  label = list(high_risk_per_hospital ~ "High risk individuals per hospital",
               hospitals ~ "Hospitals in area",
               total_at_risk ~ "Total number of high risk individuals")) |>
add_overall(col_label = "**Mean**") |>
bold_labels() |>
remove_footnote_header() |>
modify_header(label = "**Variable**") |>
modify_caption("Healthcare Capacity across metropolitain cities in New York") 
```

```{r}
#| include: false
table1$table_body
```

The mean of hospitals in the selected metropolitan area is (`r inline_text(table1, variable = "hospitals", column = "stat_0")`). There are roughly (`r inline_text(table1, variable = "total_at_risk", column = "stat_0")`) high risk individuals residing in these areas.

## **Analysis**

#### *Regression*

A regression was performed to identify the association between the number of high-risk individuals and the number of hospitals across MMSAs in New York State. Findings can be found @tbl-two.

@fig-plot displays a scatter plot between total number of high risk individuals in relation to the number of hospitals found across metropolitan areas in New York state.

```{r}
#| label: tbl-two
#| tbl-cap: "Regression Analysis"
linear <- lm(hospitals ~  total_at_risk, data = ny_mmsas)
tbl_regression(linear,intercept = TRUE,
               label = list(total_at_risk ~ "Total high risk individuals"))
```

```{r}
#| label: fig-plot 
#| fig-cap: "Scatterplot represents total number of high risk individuals in relation to the number of hospitals found across metropolitain areas in New York state."
library(ggplot2) 
library(tidyverse) 
ggplot(mmsa_icu_beds, aes(x = total_at_risk, y = hospitals)) +   
  geom_point(na.rm = TRUE, color = "royalblue3") +   
  labs(     title = "Hospitals vs. Total High-Risk Individuals",     
            x = "Total high-risk individuals",     
            y = "Number of hospitals")
```

#### *Ratio*

```{r}
#| include: false
#making new variable to then apply my function to find ratio per 100,000 people#
ny_mmsas$hospitals_per_risk_individual <- (ny_mmsas$hospitals/ ny_mmsas$total_at_risk)

hospital_ratio <- function (x)
{n <- (x*100000)
result <- (n)
return(result)
}
x<-ny_mmsas$hospitals_per_risk_individual
hospital_ratio(x)
ny_mmsas$ratio <- hospital_ratio(x)

#view and compare to verify function is correct
ny_mmsas$hospitals_per_100k_at_risk <- 
  (ny_mmsas$hospitals / ny_mmsas$total_at_risk) * 100000
view (ny_mmsas[, c("hospitals_per_100k_at_risk", "ratio")])
```

```{r}
#| label: fig-plot2
#| fig-cap: "Bar chart represents hospital ratio for every 100k high risk individual."

ggplot(ny_mmsas, aes(x =MMSA, y = ratio)) +
  geom_bar(stat = "identity", fill = "royalblue1") +
  labs(title = "Hospital Ratio for every 100k high risk individuals", x = "Metropolitain Area", y = "Ratio") +
  theme_minimal()
```

```{r}
##making statustic table
#| include: false
ny_ratio_table <- (ny_mmsas[, c("MMSA", "ratio")])
view(ny_ratio_table)
statistic <-list(
  mean=round(mean(ny_ratio_table$ratio), digits = 2),
  max=round(max(ny_ratio_table$ratio), digits = 2))
```

@fig-plot2 exhibits the ratio of hospitals to high risk individuals in the metropolitan areas in New York. There are approximately `r statistic$mean` for every 100,000 high risk individuals residing in the selected metropolitan areas in New York. The metropolitan area that has the highest amount of hospitals for every 100,000 high risk individuals in New York is Albany-Schenectady-Troy, NY, with a ratio of `r statistic$max`.
