---
title: "Healthcare Capacity in New York State"
author: Eileen Ternize
format:
  html:
    theme: yeti
execute:
  echo: false
---

## Data

The dataset combines data from the Centers for Disease Control and Prevention’s [Behavioral Risk Factor Surveillance System (BRFSS)](https://www.cdc.gov/brfss/smart/smart_2017.html), a collection of health-related surveys conducted each year of more than 400,000 Americans, and the [Kaiser Family Foundation](https://khn.org/news/as-coronavirus-spreads-widely-millions-of-older-americans-live-in-counties-with-no-icu-beds/#lookup) to show the number of people who are at high risk of becoming seriously ill from COVID-19 per ICU bed in each metropolitan area, micropolitan area or metropolitan division for which we have data.

*\*Data was found on github by jayb, covid-geography*

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(gtsummary)
library(dplyr)
library(here)

mmsa_icu_beds <- read_csv(here( "mmsa-icu-beds.csv"))
ny_mmsas <- mmsa_icu_beds %>%
  filter(str_detect(MMSA, "NY"))
```

The cleaned dataset has `r nrow(ny_mmsas)` MMSAs across New York state with complete data on total high-risk individuals and hospital counts.

## **Summary**

```{r}
#| echo: false
tbl_summary(ny_mmsas,
  by = MMSA,
  include = c(hospitals,high_risk_per_hospital, total_at_risk),
  missing = "no",
  type = list(hospitals ~ "continuous",
              high_risk_per_hospital ~ "continuous",
              total_at_risk ~ "continuous"),
statistic = all_continuous() ~ "{mean}",
  label = list(high_risk_per_hospital ~ "High risk individuals per hospital",
               hospitals ~ "Hospitals in area",
               total_at_risk ~ "Total number of high risk individuals")) |>
add_overall(col_label = "**Mean**") |>
bold_labels() |>
remove_footnote_header() |>
modify_header(label = "**Variable**") |>
modify_caption("Healthcare Capacity across metropolitain cities in New York") 
```
## **Analysis**

A regression was performed to identify the association between the number of high-risk individuals and the number of hospitals across MMSAs in New York State?

```{r}
linear <- lm(hospitals ~  total_at_risk, data = ny_mmsas)
tbl_regression(linear,intercept = TRUE,
               label = list(total_at_risk ~ "Total high risk individuals"))
```

@fig-plot displays a scatter plot between total number of high risk individuals in relation to the number of hospitals found across metropolitan areas in New York state.

```{r}
#| label: fig-plot
#| fig-cap: "Displays a scatterplot between total number of high risk individuals in relation to the number of hospitals found across metropolitain areas in New York state."
library(ggplot2)
library(tidyverse)
ggplot(mmsa_icu_beds, aes(x = total_at_risk, y = hospitals)) +
  geom_point(color = "blue") +
  labs(
    title = "Hospitals vs. Total High-Risk Individuals",
    x = "Total high-risk individuals",
    y = "Number of hospitals")
```

The ratio of hospitals to the number of high risk individuals in the selceted metropolitan areas in New York State

```{r}
#making new variable to then apply my function to find ratio per 100,000 people#
ny_mmsas$hospitals_per_risk_individual <- (ny_mmsas$hospitals/ ny_mmsas$total_at_risk)

hospital_ratio <- function (x)
{n <- (x*100000)
result <- (n)
return(result)
}

x<-ny_mmsas$hospitals_per_risk_individual
ny_mmsas$ratio <- hospital_ratio(x)

#view and compare to verify function is correct
ny_mmsas$hospitals_per_100k_at_risk <- 
  (ny_mmsas$hospitals / ny_mmsas$total_at_risk) / 100000
view (ny_mmsas[, c("hospitals_per_100k_at_risk", "ratio")])

#bar chart to see ratio
barplot(ny_mmsas$)
```