---
title: "Healthcare Capacity in New York State"
author: Eileen Ternize
format:
  html:
    theme: yeti
execute:
  echo: false
---

## Data

The dataset combines data from the Centers for Disease Control and Prevention’s [Behavioral Risk Factor Surveillance System (BRFSS)](https://www.cdc.gov/brfss/smart/smart_2017.html), a collection of health-related surveys conducted each year of more than 400,000 Americans, and the [Kaiser Family Foundation](https://khn.org/news/as-coronavirus-spreads-widely-millions-of-older-americans-live-in-counties-with-no-icu-beds/#lookup) to show the number of people who are at high risk of becoming seriously ill from COVID-19 per ICU bed in each metropolitan area, micropolitan area or metropolitan division for which we have data.

*\*Data was found on github by jayb, covid-geography*

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(gtsummary)
library(dplyr)
library(here)

mmsa_icu_beds <- read_csv(here( "mmsa-icu-beds.csv"))
ny_mmsas <- mmsa_icu_beds %>%
  filter(str_detect(MMSA, "NY"))
```

The cleaned dataset has `r nrow(ny_mmsas)` MMSAs across New York state with complete data on total high-risk individuals and hospital counts.

## **Summary**

```{r}
risk_table <- tbl_summary(ny_mmsas,
  by = MMSA,
  include = c(hospitals,high_risk_per_hospital, total_at_risk),
  missing = "no",
  type = list(hospitals ~ "continuous",
              high_risk_per_hospital ~ "continuous",
              total_at_risk ~ "continuous"),
statistic = all_continuous() ~ "{mean}",
  label = list(high_risk_per_hospital ~ "High risk individuals per hospital",
               hospitals ~ "Hospitals in area",
               total_at_risk ~ "Total number of high risk individuals")) |>
add_overall(col_label = "**Mean**") |>
bold_labels() |>
remove_footnote_header() |>
modify_header(label = "**Variable**") |>
modify_caption("Healthcare Capacity across metropolitain cities in New York") 
```

```{r}
mean_highrisk ,_ mean(ny_mmsas$hospitals)
```

```{r}
inline_text(risk_table, variable = "total_at_risk",
# 						column = "stat_0")
```

```{}
```

The average number of high-risk individuals across selected New York state metropolitan cities is `r mean_highrisk`.
